---
title: "Project_3_dataAnalysis"
author: "Project 3 group"
date: "10/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
```


```{r}
library(RJDBC)
library(DBI)


db_connection <- dbConnect(RMariaDB::MariaDB(),
                           host=Sys.getenv("TEAMTIDY_DB_HOST"),
                           user = Sys.getenv("TEAMTIDY_DB_USER"),
                           password = Sys.getenv("TEAMTIDY_DB_PASS"),
                           dbname = 'data_science_jobs')
```

```{r hard-vs-soft}
hard_vs_soft <- dbGetQuery(db_connection, "SELECT type, COUNT(*) AS n FROM skill GROUP BY type")
hard_vs_soft$n <- as.numeric(hard_vs_soft$n)
hard_vs_soft
```
```{r}
hard_vs_soft %>% ggplot(aes(x = type, y = n)) + 
  geom_bar(stat = "identity", fill = "brown", width = 0.2) + 
  ylab("Number of Jobs") + labs(title = "Hard vs Soft Skills")+ theme_bw() 
```

```{r}
#dbGetQuery(db_connection, "SELECT * FROM skill LIMIT 35")
```

```{r ds-job-skills}
job_skill <- dbGetQuery(db_connection, "SELECT description, type, COUNT(*) AS n 
                        FROM skill GROUP BY description ORDER BY n DESC")
job_skill$n <- as.numeric(job_skill$n)
head(job_skill, n =10)
```
```{r ds-job-skill-barchart}
job_skill %>% top_n(30) %>% ggplot(aes(reorder(description, n), n, fill = type)) + 
  geom_col() + coord_flip() + xlab("") + labs(title = "Top 30 Data Science job Skills") +
  theme_bw()
```



```{r hard-skill}
hard_skill <- dbGetQuery(db_connection, "SELECT description, COUNT(*) AS n 
                             FROM skill WHERE type = 'Hard Skill' GROUP BY description ORDER BY n DESC")
hard_skill$n <- as.numeric(hard_skill$n)
head(hard_skill, n = 10)
```
```{r hard-skill-barchart}
hard_skill %>% top_n(10) %>% ggplot(aes(reorder(description, n), n)) + 
  geom_col(fill = "brown") + coord_flip() + xlab("") + labs(title = "Top 10 Data Science Hard Skills") +
  theme_bw()
```

```{r soft-skill}
soft_skill <- dbGetQuery(db_connection, "SELECT description, COUNT(*) AS n 
                             FROM skill WHERE type = 'Soft Skill' GROUP BY description ORDER BY n DESC")
soft_skill$n <- as.numeric(soft_skill$n)
head(soft_skill, n =10)
```
```{r}
#dbGetQuery(db_connection, "SELECT * FROM job")
```

```{r}
data_source <- dbGetQuery(db_connection, "SELECT j.job_title, j.state, j.original_source, s.description, s.type 
           FROM job as j JOIN skill as s
           ON j.job_id = s.job_id")
head(data_source, n=10)
```
```{r}
data_source %>% count(original_source) %>% top_n(2) %>% mutate(prop = round((p = n/sum(n)),3))
```
```{r}
#data_source %>% filter(original_source == "stack overflow")
```





```{r statistics-by-type}
hard_soft_prop <- data_source %>% count(type) %>% top_n(2) %>% mutate(prop = round((p = n/sum(n)),3))
hard_soft_prop                                    
```
**Hypothesis Test:**  
Null Hypothesis, $H_{0}$: There is no difference in the proportion of hard and soft skills.$\mu_{hard} - \mu_{soft} = 0$  
Alternative Hypothesis, $H_{1}$: The proportion of hard skills is greater than proportion of soft skills.$\mu_{hard} - \mu_{soft} \geq 0$  


*Check conditions:*  
Sampling Independence: The sample is gotten from random Indeed, Stack Overflow, and LinkedIn job postings.  
Normality: Success - failure condition: np, n(1-p) > 10;  
```{r check-conditions}
prop_hard <- hard_soft_prop$prop[1] # proportion of hard skills
prop_soft <- hard_soft_prop$prop[2] # proportion of soft skills
n_hard <- hard_soft_prop$n[1] # number of hard skills
n_soft <- hard_soft_prop$n[2] # number of soft skills

#Check success - failure conditions
# Hard Skills
paste0("Hard Skills: np = ", n_hard*(prop_hard), " > 10, and n(1-p) = ",n_hard*(1 - prop_hard), " > 10")
# Soft Skills
paste0("Soft Skills: np = ", n_soft*(prop_soft), " > 10, and n(1-p) = ",n_soft*(1 - prop_soft), " > 10")
```
The Independence and success-failure condition are both satisfied. Therefore, a normal model can be assumed for this data.  


```{r hypothesis-test}
mu <- 0
alpha <- 0.05 # level of significance
df <- n_hard + n_soft - 2 # degree of freedom
diff_prop <- prop_hard - prop_soft # difference in proportion of hard and soft skills
SE <- sqrt(prop_hard*(1 - prop_hard)/n_hard + prop_soft*(1-prop_soft)/n_soft) # standard error for difference in proportions
Test_statistic <- (diff_prop - mu)/SE # Test statistic
p_value <- round(pt(Test_statistic, df, lower.tail = FALSE), 9) # p_value for one tail test
paste0("Since the p value is ", p_value, " which is less than ", alpha,
       ", we reject the null hypothesis at 0.05 level of significance.")
```
**Conclusion: ** Therefore, there is no sufficient statistical evidence that the proportion of hard skills sought for in data science job postings is equal to the proportion of soft skills sought for in data science job postings. i.e. The proportion of hard skills is greater than the proportion of soft skills.




```{r statistics-source-indeed}
source_indeed <- data_source %>% filter(original_source == "indeed") %>% count(type) %>% 
  top_n(2) %>% mutate(prop = round((p = n/sum(n)),3))
source_indeed
```

```{r}
source_indeed %>% ggplot(aes(x = type, y = prop)) + 
  geom_bar(stat = "identity", fill = "brown", width = 0.2) + 
  ylab("Proportion") + labs(title = "Indeed - Hard vs Soft Skills")+ theme_bw()
```

```{r statistics-source-stackoverflow}
source_stack <- data_source %>% filter(original_source == "stack overflow") %>% count(type) %>% top_n(2) %>% 
  mutate(prop = round((p = n/sum(n)),3))
source_stack                  
```

```{r}
source_stack %>% ggplot(aes(x = type, y = prop)) + 
  geom_bar(stat = "identity", fill = "brown", width = 0.2) + 
  ylab("Proportion") + labs(title = "Stack - Hard vs Soft Skills")+ theme_bw()
```











